Tests of measurement invariance via stochastic processes: {S}ome properties

### 2013-07-31 frontiers/frontiers.Rnw
Frontiers paper: o Added to background section o Minor update of refs.bib
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@227 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\section{Theoretical Detail}
This section contains background on the theory underlying the proposed statistics; for a more detailed account, see Merkle and Zeileis\citeyearNP{MerZei13}. 
We consider situations in which a $p$-dimensional variable $X$ with observations $x_{i},i=1,...n$ is described by a model with density $f(x_{i};\theta)$ and associated joint log-likelihood

\begin{equation}
\label{equ:likelihood}
\ell(\theta;x_{1},...,x_{n})=\overset{n}{\underset{i=1}{\Sigma}}\ell(\theta;x_{i})=\overset{n}{\underset{i=1}{\Sigma}}logf(x_{i};\theta)
\end{equation}

where $\theta$ is some $k-$dimensional parameter vector that characterizes the distribution. 
We focus on applications where the density $f(x_{i};\theta)$ arises from a structural equation model with assumed multivariate normality, though the proposed tests extend beyond this family of models. Under the usual regularity condition\cite{}, the model parameters $\theta$ can be estimated by maximum likelihood(ML),i.e.,

\begin{equation}
\label{equ:ml}
\hat{\theta}=\underset{\theta}{argmax}\ell(\theta;x_{1},...,x_{n})
\end{equation}

or equivalently by solving the first order conditions

\begin{equation}
\label{equ:max}
\overset{n}{\underset{i=1}{\Sigma}}s(\hat{\theta};x_{i})=0,
\end{equation}

where

\begin{equation}
\label{equ:score}
s(\theta;x_{i})=(\frac{\partial l(\theta;x_{i})}{\partial(\theta_{1})},...,\frac{\partial l(\theta;x_{i})}{\partial(\theta_{k})})^{T}
\end{equation}

is the score function of the model(the partial derivative of the casewise likelihood contributions w.r.t. the parameters $\theta$). Evaluation of the score function at $\hat{\theta}$ for $i=1,...n$ essentially measures the extent to which the model maximizes each individual's likelihood: as an individual's score stray further from zero,the model provides a poorer description of that individual.

>>>>

\section{Background}
This section contains background and discussion of the 
proposed statistics as applied to SEM; for a more detailed account, see
\citeA{MerZei13} and \citeA{MerFanZei}.  For details on the
statistics' application to general statistical models, see
\citeA{zeihor07}.

As currently implemented for SEM, the statistical tests described in
this paper can be applied to models that are estimated via a
multivariate normal or Wishart likelihood (or discrepancy) function
(though extension to other discrepancy functions is straightforward).
The tests are carried out after model estimation, making use of output
associated with the fitted model.  In general, we fit a model that
restricts parameters to be equal across observations, then carry out a
``posthoc'' test to examine whether certain parameters instead vary
across observations.  This procedure is similar in spirit to the
calculation of modification indices CITE and to Lagrange multiplier
tests CITE, and, in fact, these statistics can be viewed as  
special cases of the family described here.

Following model estimation, the tests primarily work on the {\em
  scores} of the fitted model; these are defined as
%% TODO indicate that this is evaluated at theta hat
\begin{equation}
\label{eq:score}
s(\hat{\theta};x_{i}) = \left ( \frac{\partial
      l(\theta;x_{i})}{\partial(\theta_{1})},...,\frac{\partial
      l(\theta;x_{i})}{\partial(\theta_{k})} \right )^{T},
\end{equation} 
where $l(\theta; x_i)$ is the likelihood associated with individual
$i$, $\theta$ is a $k$-dimensional parameter vector, and
$\hat{\theta}$ is the maximum likelihood estimate.  In words, each
individual has $k$ scores describing the extent to which the fitted
model describes that particular individual.  Scores close to zero
indicate a ``good'' description, and scores far from zero indicate a
``bad'' description.  Each of an individual's $k$ scores represent one
model parameter, so that the scores provide specific information about
each parameter's fit to each individual.

%% Next: How scores can be used for testing
